---
title: "Constrained randomization with the balancedSampling package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(balancedSampling)
```

balancedSampling helps to perform constrained randomization of a population so that the modalities of the qualitative variables are equally distributed across samples.

Lets suppose you have two variables 'A' and 'B', with four and two modalities respectively. The population is a cartesian product of these two variables (or a multiple of this cartesian product), i.e. n=8:

```{r}
data <- expand.grid(A=LETTERS[1:4], B=letters[1:2])
data
```

The frequency of the modalities are 2 (8/4) for the variable A, 4 (8/2) for variable B. There is only one partition possible, in two (since the frequency of both variables must be a multiple of the number of sample).


```{r}
balancedSampling(data, c("A", "B"), 2)
```

The samples can be further constrained by avoiding repetition of a modality:

```{r}
balancedSampling(data, c("A", "B"), 2, c(1,1))
```

Such partition is also available with only one variable, or more than two variables.

```{r}
data <- data.frame(A=rep(LETTERS[1:3], 6))
data
balancedSampling(data, "A", 3)
```

